---
title: "Chris_Final"
author: "Christopher Hoffmann"
date: "5/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading in the Data 

```{r}
library(tidyverse)
```

```{r}
top50 <- read_csv("../data/top50cleaned.csv")
glimpse(top50)
attach(top50)
```

# Variable Selection 

```{r}
library(leaps)
```

This test focuses on variable selection for the response "Popularity." In other words, finding the most optimal linear regression function for the prediction of a song's popularity.

The predictors used are:   
- Genre: Categorical variable of three categories (Hip Hop, Pop and Latin)  
- Beats Per Minute: The tempo of the song  
- Energy: Higher values indicate higher levels of energy from the song     
- Dancibilty: Higher values indicate higher dancibility of song   
- Loudness: Loudness of song measured in decibles    
- Liveness: Likelihood song is a live recording   
- Valence: The higher the value, the more positive the mood of the song  
- Length: Duration of song      
- Acousticness: The higher the value, the more acoustic the song is     
- Speechiness: Higher values indicate more lyrics  

```{r}
reg.fit <- regsubsets(Popularity ~ . -X1 -Track.Name -Artist.Name -Popularity, 
           data = top50, nvmax = 11, method = "exhaustive")

summary(reg.fit)
```

Using the Adjusted R^2 criterion, the model with the highest adjusted R^2 is the model with 7 predictors. It has an adjusted R^2 value of 0.2319. The 7 chosen predictors are:
- GenreLatin*  
- GenrePop*  
- Danceability   
- Loudness  
- Liveness  
- Valence  
- Length  

*Note: GenreLatin and GenrePop are dummy variables for the categorical variable Genre, which can either be Hip Hop, Pop or Latin.

```{r}
summary(reg.fit)$adjr2
which.max(summary(reg.fit)$adjr2)
summary(reg.fit)$adjr2[7]
```

The Mallow's Cp criterion finds that the closest Cp to p is the model with two predictors: GenrePop and Valence.
```{r}
summary(reg.fit)$cp
```

BIC chooses the thrid model, with the three predictos: GenreLatin, Valence, Length.
```{r}
summary(reg.fit)$bic
which.min(summary(reg.fit)$bic)
```

Looking at Adjusted R^2, Mallow's Cp, and the BIC, the results are all varied. However, I am going to choose the adjusted R^2 since it explains most of a song's popularity compared to the other models.

Looking at the plot we can visually compared the other models and their adjusted R^2.
```{r}
plot(reg.fit, scale = "adjr2" )
```



The coefficients for the model are below. We can see that only the intercept is very statistically significant, with GenreLatin, Valence and Length being statistically significant only if $\alpha$ < 10.
```{r}
reg.bestfit <- lm(Popularity ~ Genre + Danceability + Loudness..dB.. + Liveness + Valence. + Length., data = top50)

summary(reg.bestfit)
```

```{r}
Yhat.bestfit <- predict(reg.bestfit)
MSE.bestfit <- mean((Yhat.bestfit - Popularity)^2)
MSE.bestfit


reg.full <- lm(Popularity ~ . -X1 -Track.Name -Artist.Name -Popularity, 
               data = top50)
Yhat.full <- predict(reg.full)
MSE.full <- mean((Yhat.full - Popularity)^2)
MSE.full
```

The prediction MSE for the best fit linear model is 13.01549. The prediciton MSE for the full model was slightly smaller at 12.68, however its R^2 was 0.17 compared to 0.23. 

